Master thesis

### 2017-07-13 docs/master-thesis.tex
Thesis v1

\begin{itemize}
\tightlist
\item
  Bruno
\item
  Stéphane
\item
  Laurent
\item
  Fabian
\item
  Isabelle
\item
  Jérôme
\item
  Camilla
\item
  Aurélie
\item
  Éric
\item
  \ldots{}
\end{itemize}

@@

But several authors argue that selective logging represents a major
threat to biodiversity
\citep{Carreno-Rocabado2012, DeAvila2015, Gibson2013, Martin2015, Zimmerman2012},
challenging the sustainable definition from current selective logging.
We consequently need to assess both short and long term impacts of
selective logging on tropical forest ecosystems to implement better
syslvicultural practives in order to reach sustainability.

@@

Our work is based on the general hypothesis of a positive relationships
between biodiversity and productivity. We assumed that when a
disturbance event happen, due to a higher productivity, foerst with an
increased diversity with recover quickly and thus be more resilient.
This theoritical expectation stand on two processes: complementarity and
diversity effects \citep{Loreau2001}. If we take a species pools with
different productivity in monoculture. Their assemblage will allow theim
an overall higher productivity due to a better ressources acquisition
through ressource partitionning and niche differentiation. Additionally,
some species will individually have an higher productivity in the
assemblage than in monoculture due to facilitation from other species
present in the assemblage. Complementarity effect is the addition of the
better ressource acquisition and the facilitation. Now if we look at the
evolution of the species assemblage over time, more competitive species
will progressively dominate through competitive selection. And if
competitive species are more productive, they will increase assemblage
overall productivity. This ressources preemption by more competitive
species is the selection effect.

Generally, positive relationship between biodiversity and productivity
were emprically and experimentally demonstrated on grassland systems
\citep{Hooper2005a, Loreau2001a, Naeem2002}; but few studies focused on
the case of tropical forests. One of the few studies was realized by
\citet{Chisholm2013} and has shwon a positive significative relationship
between species richness and wood productivity on a worlwide forest
network. But those study still presents three major limits: (1) study
time are inferior to a tree life time, (2) experimental network include
scarcely disturbed plot, and (3) correlative approach does not explain
mechanisms involved in the relationship.

In the present study, we focused on mechanisms involved in the
relationship between biodiversity and forest ecosystem resilience, more
particularly tropical forest carbon resilience. We used a simulation
approach using TROLL model to assess long term processes.

@@

\begin{figure}[htbp]
\centering
\includegraphics{images/TROLLtree.png}
\caption{\label{fig:TROLLtree}Individuals tree inside TROLL explicit spatial
grid from \citet{Li}. Tree geometry (crown radius CR, crown depth CD,
height h, diameter at breast height dbh) is updated at each timestep
following allometric relationship with assimilated carbon allocated to
growth. Each tree is flagged with a species label linking to its
species-specific attributes. Light is compputed explicitly at each
timestep for each voxel.}
\end{figure}

@@

\(Death\) of a tree follows a \(\mathcal{Bernoulli}\) law of probability
\(P(Death)\). The odds for a tree to die are calculated with the sum of
the natural tree death odd \(\theta\) and a perturbation index
\(\beta*e^{\alpha*d_{gaps}}\). The perturbation index depend on the
distance \(d_{gaps}\) of the tree \(i\) to the closest logging gap. The
probability for a tree to die \(P(Death)\) is finally calculated by
taking the inverse logit \(logit^{-1}\) of the odd.

@@

\citet{Li} already assessed TROLL model sensitivity to several
parameters (\(k\) see \eqref{eq:PPFD}, \(\phi\) see \eqref{eq:LCP}, \(g1\)
see \eqref{eq:gs}, \(f_{wood}\) see \eqref{eq:DeltaV}, \(f_{canopy}\) see
\eqref{eq:DeltaLA} and \(m\) see \eqref{eq:db}) which they assumed having a
key role in model functioning. On the other hand, we decided to use
TROLL to study resistance and resilience of ecosystem face to
disturbance, highlighting the role of biodiversity. Consequently we
particulrarly needed to assess the importance of functional traits to
further better control and evaluate functional diversities. We also
needed to assess the sensitivity of TROLL model to the seed rain
constant (\(n_{ext}\), see \eqref{eq:next}) because we assumed it as one
of the main factors of tree recruitments after disturbance within
simulations.

@@

Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and
188.9 GB of memory. We assumed maturity of the forest after 500 years of
regeneration (Maréchaux \& Chave) and computed simulation 100 years
after a disturbance event with 40\% loss of basal area. Due to computer
limitations we did not run replicate (besides it should be necessary to
reduce simulation stochasticity). To assess ecosystem outputs
sensitivity to studied parameters, we compared it to 100 replicates of
control simulations with all parameters set to default values. Ecosystem
outputs outside of the range of the control replicates values are
significantly influenced by the studied parameter.

@@

In order to assess the role of biodiversity in ecosystem answer to both
disturbance and sylviculture, we needed to create a space of experiments
encompassing both variation of disturbance, biodiversity and time.
Disturbance was represented by percentage of basal area loss (0\%, 25\%,
50\% and 75\%), or as a selective logging simulation using default
parameters. Biodiversity was integrated with two components of its
components: taxonomic and functional diversities. We used species
richness \(SR\) to represents taxonomic diversity (5, 25, and 125
species). Functional diversity can be related to numerous components,
and \citet{Borgy2017} argued for 5: richness, divergence, regularity,
overlap and mean. Because mature forest were created from a bare soil
with TROLL simulations, we could not control a priori divergence,
regularity and overlap but only assess theim before diturbance.
Consequently, we focused on functional richness with convex hull volume
\(CHV\) and functional mean with community weighted mean \(CWM\). For
each level of species richness \(SR\), we selected 20 communities with
growing convex hull volume \(CHV\) but with a community weighted means
close to the regional species pool community weighted means.
Effectivelly, we did not wanted drastic change in community means that
could have more effect than functional richness itself. This design of
experiments resulted in 60 communities (\(5~SR*20~CHV\)) and 240
simulations (\(60 ~communities*4~levels~of~disturbance\)) over 600 years
(maturity being assumed after 500 years of regeneration \citep{Li}).
Figure \ref{fig:DOE} presents the design of experiment for communities
biodiversity after the mature forest were simulated, and thus before
disturbance. We obtained a broad range of both functionl dispersion
\(FDis\) and aboveground biomass \(AGB\) for simulated forest ecosystems
before disturbance.

@@

The resilience of the system \(R(t)\) at the time \(t\) is described by
the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
\(Loss(t_d)\) at disturbance time \(t_d\). But in our peculiar case of
tropical forest ecosystems, the equilibrium used to calculate
\(Loss(t_d)\) can not be reduced to a specific time if the equilibrium
is dynamic. Consequently, to encompass undisturbed ecosystem variations
throught time, we simulated an undisturbed control ecosystem \(C\). And
the resilience of the system \(R(t)\) at the time \(t\) was defined as
the ratio of the ecosystem metric values in the disturbed simulation
\(X_T(t)\) over the the ecosystem metric values from the control
\(X_C(t)\) . Thus, the value of resilience \(R(t)\) is normalized for
all simulations and metrics. \(R(t)\) will be equal to \(R_{eq} = 1\)
when reaching the equilibrium value. Consequently we can calculate an
euclidean distance to equilibrium \(d_{eq}(t)\) as
\(d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}\). Ecosystem euclidean distance
to equilibrium was calculated in a multi-dimensional space for the two
functions described above: forest dynamic (AGB, BA, N, N10, and N30) and
forest production (GPP, NPP, Rday, and Rnight). We then used cummulative
integral over time of eulcidean distance to equilibrium to assess
simulations resilience.

@@

Most of functional traits had a significant long term influence on
ecosystem outputs (figure \ref{fig:ftVar}). On the other hand, few
functional traits influenced final ecosystem structure. Only specific
maximum diameter \textbf{dmax} add higher diversity for greater orders
implying better eveness in species distributions. Regarding functional
composition, traits fixed to mean did not change other functional traits
density distribution. Moreover, seedrain did not seem to affect
aboveground biomass and final ecosystem height and diameter structure.
Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a
similar effect. Lower seedrain implied faster decrease of stem above 10
cm dbh and higher number of stem above 30 cm diameter at breast height
after ecosystem resilience to disturbance (approximately 50 years).
Lower seedrain than default decreased basal area over time. In addition,
lower seedrain than default decreased equitability by increasing
abundance of abundant species and decreasing abundance of less abundant
species. See
\protect\hyperlink{appendix-4-sensitivity-analysis}{Appendix 4:
Sensitivity analysis} for further details.

@@

The ranking was stable over time for the 240 simulations. So we used the
cummulative integral after 600 years \(Ieq_{600}\) as a measurement of
ecosystem resilience. We compared cummulative integral after 600 years
to communities taxonomic and functional diversity for each leavel of
disturbance (see Figure \ref{fig:gIeq}). We found that increased
functional diversity \citep[FDiv,][]{villeger_new_2008} was reducing
cummulative integral from ecosystem distance to forest dynamic
equilibrium after 600 years (\(Ieq_{600}\)). In addition, functional
eveness was complementary reducing \(Ieq_{600}\). Finally species
richness was not directly link to \(Ieq_{600}\). Effectivelly, low
species richness could result in variant \(Ieq_{600}\), but increased
species richness resulted in increased functional diversity and
consequently lower \(Ieq_{600}\). We found similar results for all
disturbance levels and other ecosystem functions (forest production and
forest diversity, see
\protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
Disturbance simulations}).

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/gIeq-1.pdf}
\caption{\label{fig:gIeq}Ecosystem resilience after 600 years with taxonomic
and functional diversity for different levels of disturbance.
Cummulative integral from ecosystem distance to forest dynamic
equilibrium after 600 years was represented against functional diversity
\citep[FDiv,][]{villeger_new_2008} for different level of disturbance
(25, 50 and 75\% of total basal area); dot shapes represents the species
richness whereas dot color represents functional eveness
\citep[FEve,][]{villeger_new_2008}.}
\end{figure}

@@

In the limit of the model, we were able to show that diversity improved
tropical forest resilience. More particularly, functional diversity and
evenness are key components of diversity in forest recovery after
disturbance. Moreover, we found that complementarity between species was
insuring forest recovery in forest succession start with facilitation
before more productive species dominate the forest and insure recovery.
Our results advocates for a sustainable harvesting of tropical forests
through an increased resilience due to high diversity. But this
conclusion should met a sustainable definiton of selective logging
following \citet{Zimmerman2012}. Because if the harvesting are not
sustainable negative feedbaks will slowly diminish diversity and its
benefits for forest resilience, resulting in forest degradation.

@@

Our results validated the hypothesis of a significative relationship
between forest resilience and functional diversity and eveness
(\(p < 0.01\) and \(p < 0.05\) respectivelly). Thus we were able to show
that diversity improve forest resilience. More particularly, functional
diversity seems a major aspect of resilience if its enforced by a high
functional eveness. Effectively high functional diversity needs eveness
in order to better answer disturbance, if not the diversity is masked by
ecosystem dominant species. Those results confirms the review of
\citet{Diaz2001} advocating for underevaluated importance of plant
functional diversity in ecosystem processes. Additionally, the role of
eveness confirm the review of \citet{Zhang2012} who highlighted the role
of eveness in productivity and thus resilience following our hypothesis.
Finally, species and functional richnesses is not directly increasing
resilience. But increased species and functional richnesses will
increase chance for high functional diversity through the sampling
effect \citep{Loreau1998}. An higher sampling of regional species pool
will allow a greater chance to pick more functionally diverse species.

@@

We found that complementary effect was insuring forest resilience in the
beginning of forest successions. We interpreted this results as the
consequence of facilitation processes. As the only ressource simulated
by TROLL is the light, the major facilitation will be light shading of
post-pioneer species by pioneer species in disturbance gaps. But the
complementarity effect is reducing through time, to let the selection
effect insure forest resilience due to an inforced dominance of more
productive species. The diminishing of complementarity effect is due to
competitive selection though time in gaps succession. But the study
scale matters as shwon by \citet{Chisholm2013}. Here we look at
processes to a 16 ha scale. We do not have topography reducing
micro-environment effects, but \citet{Chisholm2013} results suggest that
complementarity will be stronger at smaller scale and could explain its
low value after forest recovery. Finally, our findings confirm results
of \citet{Tobner2016} realized on experimental forests of 4 years in low
diverse forest of Canada. They also found that selection effect was
greater than complementarity in most of the case. In the case of
selective logging the complementarity effect will thus be the major
effect between two cutting cycles due to the cycle length (several
decades in guyana shield). High forest diversity with important species
complementarity through increased functional diversity is thus an
advantage for forest recovery between two rotation in order to maintain
both productive ecosystem and sustainable management.

@@

We used closed forest system simulated with TROLL forest model to
evaluate the role and mechanisms of biodiversity in tropical forest
resilience to disturbance. We found that diversity improved forest
resilience and that complementarity between species was insuring forest
recovery in forest succession start with facilitation before more
productive species dominate the forest and insure recovery. Our results
advocates for sustainable selective logging against monoculture stand:
high diversity will increase forest resilience and thus improve logging
cycles. Even if monoculture stand are naturalized they will not reach
mature forest diversity, and consequently they will not reach its
natural high resilience. But on the other hand selective logging in
tropical forest needs to be sustainable, if not the diversity will
slowly decrease after each cycles degrading forest ecosystem functions,
and resilience will decrease due to negative feedbacks.
\citet{Zimmerman2012} criticized the state of sylviculture in tropics,
suggesting that ``we have not been able to reconcile these opposing
biological and economic forces''. Still, they advocates for a
possibility of sustainable tropical logging, already existing in
small-scale, that will need proper funding from the internationnal
communtiy. This view correspond to the high resilience, we were able to
show, of tropical forest hyperdiverse systems and let hope for a future
sustainable selective logging in tropical forests.

@@

We compiled functional traits from GLOPNET
\citep{wright_worldwide_2004}, TRY \citep{Kattge2011}, and DRYAD
\citep{chave_towards_2009} databases (see \ref{tab:A2traits}). We kept
dataset given by GLOPNET as origin dataset for observations. Dataset
defined as origin corresponded to leaf lifespan (\(LL\)) and most of the
time to leaf mass per area (\(LMA\)) and leaf nitrogen content per leaf
dry mass (\(Nmass\)). We measured variable importance in functionnal
traits to explain leaf lifespan with out-of bag method applied on a
random forest. Then, we used a béayesian apporach to test different
models with growing level of complexity. We retained the model with the
best tradeoff between model complexity (number of parameters \(K\)),
convergence, likelihood, and prediction quality (root mean sqaure error
of prediciton \(RMSEP\)). We finally tested the new allometry obtained
with the selected model with TROLL simulations.

@@

Simulations are validating that this new allometry resolve the issue of
unrealistically low aboveground biomass for low LMA species due to an
early edath of individuals inside simulations. For instance with this
allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic
aboveground biomass above 400 \(tonC.ha^{-1}\) and realistic diameter
and age distribution inside the final population.

@@

\[V_f \sim \mathcal{logN}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]\]

@@

\begin{longtable}[]{@{}ll@{}}
\caption{\label{tab:A3models}Models summary.}\tabularnewline
\toprule
M & Model\tabularnewline
\midrule
\endfirsthead
\toprule
M & Model\tabularnewline
\midrule
\endhead
\(M_{s,p}\) &
\({P_{rotten}}_n \sim \mathcal{B}(inv_{logit}(\beta_0 + \beta_1*dbh_n + {\beta_2}_p + {\beta_3}_s))\)\tabularnewline
\(N_{s,p} + L_{s,p}\) &
\(Volume_{of~wood} \sim \mathcal{logN} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)\tabularnewline
\bottomrule
\end{longtable}

@@

\(N_{s,p} + L_{s,p}\):
\(Volume_{of~wood} \sim \mathcal{logN} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)

@@

To study resistance and resilience of ecosystem face to disturbance,
highlighting the role of biodiversity, we decided to use TROLL model
simulations \citep{Chave1999}. In order to get a finer study of
simulations response, we need to assess first sensitivity of the TROLL
model to different parameters. More particularly we need to assess the
importance of functional traits to further better control functional
diversities in simulations. We also need to assess sensitivity of the
model to see rain constant because we assume it is one of the main
factor of tree recruitments after disturbance in the model.

@@

To assess the sensitivity of TROLL model to species functionnal traits,
we performed a sensitivity analysis by fixing species trait values to
their mean. Each trait was tested independently. We reduce to a common
mean traits with a correlation \(r \geq 0.8\) (see figure
\ref{fig:A4corr}).

@@

Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and
188.9 GB of memory. We assumed maturity of the forest after 500 years of
regeneration \citep{Li} and computed simulation 100 years after a
disturbance event of 40\% intensity. Due to computer limitations we did
not run replicate (besides it should be necessary to reduce simulation
stochasticity). To assess ecosystem outputs sensitivity to studied
parameters, we compared it to 100 replicates of control simulations with
all parameters set to default values. Ecosystem outputs outside of the
range of the control replicates values are significantly influenced by
the studied parameter.

>>>>

\begin{itemize}
\tightlist
\item
  Bruno
\item
  Stéphane
\item
  Laurent
\item
  Fabian
\item
  Isabelle
\item
  Jérôme
\item
  Camilla
\item
  Aurélie
\item
  Éric
\item
  Maxime
\item
  Raphaël
\item
  Paracou team
\item
  \ldots{}
\end{itemize}

@@

But several authors argue that selective logging represents a major
threat to biodiversity
\citep{Carreno-Rocabado2012, DeAvila2015, Gibson2013, Martin2015, Zimmerman2012},
challenging the sustainable definition from current selective logging.
We consequently need to assess both short and long term impacts of
selective logging on tropical forest ecosystems to implement better
sylvicultural practices in order to reach sustainability.

@@

Our work is based on the general hypothesis of a positive relationships
between biodiversity and productivity. We assumed that when a
disturbance event happen, due to a higher productivity, forest with an
increased diversity will recover quicker and thus be more resilient.
This theoritical expectation stand on two processes: complementarity and
diversity effects \citep{Loreau2001}. If we take a species pools with
different productivity in monoculture. Their assemblage will allow theim
an overall higher productivity due to a better ressources acquisition
through ressource partitionning and niche differentiation. Additionally,
some species will individually have an higher productivity in the
assemblage than in monoculture due to facilitation from other species
present in the assemblage. Complementarity effect is the addition of the
better ressource acquisition and the facilitation. Now if we look at the
evolution of the species assemblage over time, more competitive species
will progressively dominate through competitive selection. And if
competitive species are more productive, they will increase assemblage
overall productivity. This ressources preemption by more competitive
species is the selection effect.

Generally, positive relationship between biodiversity and productivity
were emprically and experimentally demonstrated on grassland systems
\citep{Hooper2005a, Loreau2001a, Naeem2002}; but few studies focused on
the case of tropical forests. One of the few studies was realized by
\citet{Chisholm2013} and has shown a positive significative relationship
between species richness and wood productivity on a worlwide forest
network. But those study still presents three major limits: (1) study
time are inferior to a tree life time, (2) experimental network include
scarcely disturbed plot, and (3) correlative approach does not explain
mechanisms involved in the relationship.

In the present study, we focused on mechanisms involved in the
relationship between biodiversity and forest ecosystem resilience. We
used a simulation approach using TROLL model to assess long term
processes.

@@

\begin{figure}[htbp]
\centering
\includegraphics{images/TROLLtree.png}
\caption{\label{fig:TROLLtree}Individuals tree inside TROLL explicit spatial
grid from \citet{Li}. Tree geometry (crown radius CR, crown depth CD,
height h, diameter at breast height dbh) is updated at each timestep
following allometric relationship with assimilated carbon allocated to
growth. Each tree is flagged with a species label linking to its
species-specific attributes. Light is computed explicitly at each
timestep for each voxel.}
\end{figure}

@@

\(Death\) of a tree follows a \(\mathcal{B}ernoulli\) law of probability
\(P(Death)\). The odds for a tree to die are calculated with the sum of
the natural tree death odd \(\theta\) and a perturbation index
\(\beta*e^{\alpha*d_{gaps}}\). The perturbation index depend on the
distance \(d_{gaps}\) of the tree \(i\) to the closest logging gap. The
probability for a tree to die \(P(Death)\) is finally calculated by
taking the inverse logit \(logit^{-1}\) of the odd.

@@

\citet{Li} already assessed TROLL model sensitivity to several
parameters (\(k\) see \eqref{eq:PPFD}, \(\phi\) see \eqref{eq:LCP}, \(g1\)
see \eqref{eq:gs}, \(f_{wood}\) see \eqref{eq:DeltaV}, \(f_{canopy}\) see
\eqref{eq:DeltaLA} and \(m\) see \eqref{eq:db}) which they assumed having a
key role in model functioning. On the other hand, we decided to use
TROLL to study resistance and resilience of ecosystem face to
disturbance, highlighting the role of biodiversity. Consequently we
particulrarly needed to assess the importance of functional traits to
further better control and evaluate functional diversities. We also
needed to assess the sensitivity of TROLL model to the seed rain
constant (\(n_{ext}\), see \eqref{eq:next}) because we assumed it was one
of the main factors of tree recruitments after disturbance within
simulations.

@@

Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and
188.9 GB of memory. We assumed maturity of the forest after 500 years of
regeneration \citet{Li} and computed simulation 100 years after a
disturbance event with 40\% loss of basal area. Due to computer
limitations we did not run replicates (besides it should be necessary to
reduce simulation stochasticity). To assess ecosystem outputs
sensitivity to studied parameters, we compared it to 100 replicates of
control simulations with all parameters set to default values. Ecosystem
outputs outside of the range of the control replicates values are
significantly influenced by the studied parameter.

@@

In order to assess the role of biodiversity in ecosystem answer to both
disturbance and sylviculture, we needed to create a space of experiments
encompassing both variation of disturbance, biodiversity and time.
Disturbance was represented by percentage of basal area loss (0\%, 25\%,
50\% and 75\%), or as a selective logging simulation using default
parameters. Biodiversity was integrated with two of its components:
taxonomic and functional diversities. We used species richness \(SR\) to
represents taxonomic diversity (5, 25, and 125 species). Functional
diversity can be related to numerous components, and \citet{Borgy2017}
argued for 5: richness, divergence, regularity, overlap and mean.
Because mature forest were created from a bare soil with TROLL
simulations, we could not control a priori divergence, regularity and
overlap but only assess theim before diturbance. Consequently, we
focused on functional richness with convex hull volume \(CHV\) and
functional mean with community weighted mean \(CWM\). For each level of
species richness \(SR\), we selected 20 communities with growing convex
hull volume \(CHV\) but with a community weighted means close to the
regional species pool community weighted means. Effectivelly, we did not
wanted drastic change in community means that could have more effect
than functional richness itself. This design of experiments resulted in
60 communities (\(5~SR*20~CHV\)) and 240 simulations
(\(60 ~communities*4~levels~of~disturbance\)) over 600 years (maturity
being assumed after 500 years of regeneration \citep{Li}). Figure
\ref{fig:DOE} presents the design of experiment for communities
biodiversity after the mature forest were simulated, and thus before
disturbance. We obtained a broad range of both functionl dispersion
\(FDis\) and aboveground biomass \(AGB\) for simulated forest ecosystems
before disturbance.

@@

The resilience of the system \(R(t)\) at the time \(t\) is described by
the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
\(Loss(t_d)\) at disturbance time \(t_d\). But in our peculiar case of
tropical forest ecosystems, the equilibrium used to calculate
\(Loss(t_d)\) can not be reduced to a specific time if the equilibrium
is dynamic. Consequently, to encompass undisturbed ecosystem variations
throught time, we simulated an undisturbed control ecosystem \(C\). And
the resilience of the system \(R(t)\) at the time \(t\) was defined as
the ratio of the ecosystem metric values in the disturbed simulation
\(X_T(t)\) over the ecosystem metric values from the control \(X_C(t)\)
. Thus, the value of resilience \(R(t)\) is normalized for all
simulations and metrics. \(R(t)\) will be equal to \(R_{eq} = 1\) when
reaching the equilibrium value. Consequently we can calculate an
euclidean distance to equilibrium \(d_{eq}(t)\) as
\(d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}\). Ecosystem euclidean distance
to equilibrium was calculated in a multi-dimensional space for the two
functions described above: forest dynamic (AGB, BA, N, N10, and N30) and
forest production (GPP, NPP, Rday, and Rnight). We then used cummulative
integral over time of eulcidean distance to equilibrium to assess
simulations resilience.

@@

Most of functional traits had a significant long term influence on
ecosystem outputs (figure \ref{fig:ftVar}). On the other hand, few
functional traits influenced final ecosystem structure. Only specific
maximum diameter \textbf{dmax} add higher diversity for greater orders
implying better evenness in species distributions. Regarding functional
composition, traits fixed to mean did not change other functional traits
density distribution. Moreover, seedrain did not seem to affect
aboveground biomass and final ecosystem height and diameter structure.
Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a
similar effect. Lower seedrain implied faster decrease of stem above 10
cm dbh and higher number of stem above 30 cm diameter at breast height
after ecosystem resilience to disturbance (approximately 50 years).
Lower seedrain than default decreased basal area over time. In addition,
lower seedrain than default decreased equitability by increasing
abundance of abundant species and decreasing abundance of less abundant
species. See
\protect\hyperlink{appendix-4-sensitivity-analysis}{Appendix 4:
Sensitivity analysis} for further details.

@@

The ranking was stable over time for the 240 simulations. So we used the
cummulative integral after 600 years \(Ieq_{600}\) as a measurement of
ecosystem resilience. We compared cummulative integral after 600 years
to communities taxonomic and functional diversity for each leavel of
disturbance (see Figure \ref{fig:gIeq}). We found that increased
functional diversity \citep[FDiv,][]{villeger_new_2008} was reducing
cummulative integral from ecosystem distance to forest dynamic
equilibrium after 600 years (\(Ieq_{600}\)). In addition, functional
evenness was complementary reducing \(Ieq_{600}\). Finally species
richness was not directly link to \(Ieq_{600}\). Effectivelly, low
species richness could result in variant \(Ieq_{600}\), but increased
species richness resulted in increased functional diversity and
consequently lower \(Ieq_{600}\). We found similar results for all
disturbance levels and other ecosystem functions (forest production and
forest diversity, see
\protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
Disturbance simulations}).

\begin{figure}[htbp]
\centering
\includegraphics{master-thesis_files/figure-latex/gIeq-1.pdf}
\caption{\label{fig:gIeq}Ecosystem resilience after 600 years with taxonomic
and functional diversity for different levels of disturbance.
Cummulative integral from ecosystem distance to forest dynamic
equilibrium after 600 years was represented against functional diversity
\citep[FDiv,][]{villeger_new_2008} for different level of disturbance
(25, 50 and 75\% of total basal area); dot shapes represents the species
richness whereas dot color represents functional evenness
\citep[FEve,][]{villeger_new_2008}.}
\end{figure}

@@

In the limit of the model, we were able to show that diversity improved
tropical forest resilience. More particularly, functional diversity and
evenness are key components of diversity in forest recovery after
disturbance. Moreover, we found that complementarity between species was
insuring forest recovery in forest succession start with facilitation
before more productive species dominate the forest and insure recovery.
Our results advocates for a sustainable harvesting of tropical forests
through an increased resilience due to high diversity. But this
conclusion should met a sustainable definiton of selective logging
following \citet{Zimmerman2012}. Because if the harvesting is not
sustainable negative feedbaks will slowly diminish diversity and its
benefits for forest resilience, resulting in forest degradation.

@@

Our results validated the hypothesis of a significative relationship
between forest resilience and functional diversity and evenness
(\(p < 0.01\) and \(p < 0.05\) respectivelly). Thus we were able to show
that diversity improve forest resilience. More particularly, functional
diversity seems a major aspect of resilience if its enforced by a high
functional evenness. Effectively high functional diversity needs
evenness in order to better answer disturbance, if not the diversity is
masked by ecosystem dominant species. Those results confirms the review
of \citet{Diaz2001} advocating for underevaluated importance of plant
functional diversity in ecosystem processes. Additionally, the role of
evenness confirm the review of \citet{Zhang2012} who highlighted the
role of evenness in productivity and thus resilience following our
hypothesis. Finally, species and functional richnesses is not directly
increasing resilience. But increased species and functional richnesses
will increase chance for high functional diversity through the sampling
effect \citep{Loreau1998}. An higher sampling of regional species pool
will allow a greater chance to pick more functionally diverse species.

@@

We found that complementary effect was insuring forest resilience in the
beginning of forest successions. We interpreted this results as the
consequence of facilitation processes. As the only ressource simulated
by TROLL is the light, the main facilitation will be light shading of
post-pioneer species by pioneer species in disturbance gaps. But the
complementarity effect is reducing through time, to let the selection
effect insure forest resilience due to an inforced dominance of more
productive species. The diminishing of complementarity effect is due to
competitive selection though time in gaps succession. But the study
scale matters as shwon by \citet{Chisholm2013}. Here we look at
processes to a 16 ha scale. We do not have topography reducing
micro-environment effects, but \citet{Chisholm2013} results suggest that
complementarity will be stronger at smaller scale and could explain its
low value after forest recovery. Finally, our findings confirm results
of \citet{Tobner2016} realized on experimental forests of 4 years in low
diverse forest of Canada. They also found that selection effect was
greater than complementarity in most of the cases. In the case of
selective logging the complementarity effect will thus be the major
effect between two cutting cycles due to the cycle length (several
decades in guyana shield). High forest diversity with important species
complementarity through increased functional diversity is thus an
advantage for forest recovery between two cutting cycles in order to
maintain both productive ecosystem and sustainable management.

@@

We used closed forest system simulated with TROLL forest model to
evaluate the role and mechanisms of biodiversity in tropical forest
resilience to disturbance. We found that diversity improved forest
resilience. Additionally we found that complementarity between species
was insuring forest recovery in forest succession start with
facilitation, before more productive species dominate the forest and
insure forest recovery. Our results advocate for sustainable selective
logging against monoculture stand: high diversity will increase forest
resilience and thus improve logging cycles. Even if monoculture stand
are naturalized they will not reach mature forest diversity, and
consequently they will not reach its natural high resilience. But on the
other hand selective logging in tropical forest needs to be sustainable,
if not the diversity will slowly decrease after each cycles degrading
forest ecosystem functions, and resilience will decrease due to negative
feedbacks. \citet{Zimmerman2012} criticized the state of sylviculture in
tropics, suggesting that ``we have not been able to reconcile these
opposing biological and economic forces''. Still, they advocates for a
possibility of sustainable tropical logging, already existing in
small-scale, that will need proper funding from the internationnal
communtiy. This view correspond to the high resilience of tropical
forest hyperdiverse systems, we were able to show, and let hope for a
future sustainable selective logging in tropical forests.

@@

We compiled functional traits from GLOPNET
\citep{wright_worldwide_2004}, TRY \citep{Kattge2011}, and DRYAD
\citep{chave_towards_2009} databases (see \ref{tab:A2traits}). We kept
dataset given by GLOPNET as origin dataset for observations. Dataset
defined as origin corresponded to leaf lifespan (\(LL\)) and most of the
time to leaf mass per area (\(LMA\)) and leaf nitrogen content per leaf
dry mass (\(Nmass\)). We measured variable importance in functionnal
traits to explain leaf lifespan with an out-of the bag method applied on
a random forest. Then, we used a bayesian apporach to test different
models with growing level of complexity. We retained the model with the
best tradeoff between model complexity (number of parameters \(K\)),
convergence, likelihood, and prediction quality (root mean sqaure error
of prediciton \(RMSEP\)). We finally tested the new allometry obtained
with the selected model with TROLL simulations.

@@

Simulations are validating that this new allometry resolve the issue of
unrealistically low aboveground biomass for low LMA species due to an
early death of individuals inside simulations. For instance with this
allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic
aboveground biomass above 400 \(tonC.ha^{-1}\) and realistic diameter
and age distribution inside the final population.

@@

\[V_f \sim log\mathcal{N}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]\]

@@

\begin{longtable}[]{@{}ll@{}}
\caption{\label{tab:A3models}Models summary.}\tabularnewline
\toprule
M & Model\tabularnewline
\midrule
\endfirsthead
\toprule
M & Model\tabularnewline
\midrule
\endhead
\(M_{s,p}\) &
\({P_{rotten}}_n \sim \mathcal{B}(inv_{logit}(\beta_0 + \beta_1*dbh_n + {\beta_2}_p + {\beta_3}_s))\)\tabularnewline
\(N_{s,p} + L_{s,p}\) &
\(Volume_{of~wood} \sim log\mathcal{N} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)\tabularnewline
\bottomrule
\end{longtable}

@@

\(N_{s,p} + L_{s,p}\):
\(Volume_{of~wood} \sim log\mathcal{N} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)

@@

To study resistance and resilience of ecosystem face to disturbance,
highlighting the role of biodiversity, we decided to use TROLL model
simulations \citep{Chave1999}. In order to get a finer study of
simulations response, we needed to assess sensitivity of the TROLL model
to different parameters. More particularly we needed to assess the
importance of functional traits to further better control functional
diversities in simulations. We also needed to assess sensitivity of the
model to see rain constant because we assumed it was one of the main
factor of tree recruitments after disturbance in the model.

@@

To assess the sensitivity of TROLL model to species functionnal traits,
we performed a sensitivity analysis by fixing species trait values to
their mean. Each trait was tested independently. We reduced to a common
mean traits with a correlation \(r \geq 0.8\) (see figure
\ref{fig:A4corr}).

@@

Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and
188.9 GB of memory. We assumed maturity of the forest after 500 years of
regeneration \citep{Li} and computed simulation 100 years after a
disturbance event of 40\% intensity. Due to computer limitations we did
not run replicates (besides it should be necessary to reduce simulation
stochasticity). To assess ecosystem outputs sensitivity to studied
parameters, we compared it to 100 replicates of control simulations with
all parameters set to default values. Ecosystem outputs outside of the
range of the control replicates values are significantly influenced by
the studied parameter.
